### 2020-04-23 dataclean.tex
final tweaks

% !TEX root = main.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Viewing and Cleaning the Data
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[bt]
  \includegraphics[width=6.5in]{figs/sample_sn.pdf}
  \caption{Serial number view of GPU life and failures. Legend: black
    dots are installs, black lines are lifetimes at installed
    location, blue squares are OTB events, red triangles are DBE
    events, and black ] are ``last seen'' events.}
  \label{fig:gpuview}
\end{figure*}
\section{Viewing and Cleaning the Data}
\label{section:dataclean}
As we need to recover durations of GPU operation from this data, correct
processing involves time adjustments for switching between daylight
saving time and standard time and leap years. We perform this by
setting a reference time zone (Eastern time) and converting all
date-times from strings into POSIX date-time variables with the R
\pkg{lubridate} package \cite{lubridate}, which enables appropriate
date arithmetic and sensible date constructs for graphs.

@@

To get some intuition for the GPU lifetimes on Titan, we give two
views of 80 randomly selected GPUs (Figure~\ref{fig:gpuview}) and 80
randomly selected locations (Figure~\ref{fig:locview}).
\begin{figure*}[tb]
  \includegraphics[width=6.5in]{figs/sample_loc.pdf}
  \caption{Location view of GPU life and failures.  Legend: black dots
    are installs, black lines are lifetimes at installed location,
    blue squares are OTB events, red triangles are DBE events, and
    black ] are ``last seen'' events.}
  \label{fig:locview}
\end{figure*}
The GPU view
visually documents the life of each GPU unit: when it was installed and
removed at various locations, its DBE and OTB events, and the last
time it was seen. The location view documents the life of a location:
when different GPUs were installed and removed, their OTB and DBE
events, and whether a removal was the last time the unit was seen on
the system. These views were critical to understanding the data and to
verifying various data processing decisions.

>>>>

% !TEX root = main.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Viewing and Cleaning the Data
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[bt]
  \includegraphics[width=6.5in]{figs/sample_sn.pdf}
  \caption{Serial number view of GPU life and failures. Legend: black
    dots are installs, black lines are lifetimes at installed
    location, blue squares are OTB events, red triangles are DBE
    events, and black ] are ``last seen'' events.}
  \label{fig:gpuview}
\end{figure*}
\begin{figure*}[tb]
  \includegraphics[width=6.5in]{figs/sample_loc.pdf}
  \caption{Location view of GPU life and failures.  Legend: black dots
    are installs, black lines are lifetimes at installed location,
    blue squares are OTB events, red triangles are DBE events, and
    black ] are ``last seen'' events.}
  \label{fig:locview}
\end{figure*}
\section{Viewing and Cleaning the Data}
\label{section:dataclean}
As we need to recover durations of GPU operation from this data, correct
processing involves time adjustments for switching between daylight
saving time and standard time and leap years. We perform this by
setting a reference time zone (Eastern time) and converting all
date-times from strings into POSIX date-time variables with the R
\pkg{lubridate} package \cite{lubridate}, which enables appropriate
date arithmetic and sensible date constructs for graphs.

@@

To get some intuition for the GPU lifetimes on Titan, we give two
views of 60 randomly selected GPUs (Figure~\ref{fig:gpuview}) and 80
randomly selected locations (Figure~\ref{fig:locview}).
The GPU view
visually documents the life of each GPU unit: when it was installed and
removed at various locations, its DBE and OTB events, and the last
time it was seen. The location view documents the life of a location:
when different GPUs were installed and removed, their OTB and DBE
events, and whether a removal was the last time the unit was seen on
the system. These views were critical to understanding the data and to
verifying various data processing decisions.

### 2020-04-23 discussion.tex
Fixed related work and discussion section.
final tweaks

\fix{Possibly a short summary of the observations in this paper to facilitate the discussion below.}

@@

The options of replacing failed components and employing reliability-aware
resource management may not be available or cost effective for other
supercomputing centers dealing with similar unexpected reliability issues.
%
Replacement components may not be readily available and may have to be
manufactured, which can be impossible for technology that is no longer supported
by the original manufacturer. Experienced errors or failures may be caused by
software, which can be difficult to fix if the software itself or the deployed
version is no longer supported by the vendor.

There is also a financial aspect, such as service contract and/or warranty
expiration and the operational budget of a supercomputing center not having
significant funds for replacement/reengineering costs or a manufacturer or
vendor not paying for them.
%
There is additionally the time and financial cost for root-cause analysis and
developing a hardware and/or software mitigation strategy in the first place.
Fixing a problem requires finding and understanding it, which can be difficult 
in today's complex systems and may require knowledge that only the original 
manufacturer or vendor possesses.

Reliability-aware resource scheduling also has its limitations, as the network
architecture needs to be taken into account. Titan's 3D torus network creates
certainly more challenges for efficient job allocations, than the fat tree
network of Titan's successor, Summit~\cite{olcf:summit}. Nodes associated with
the same job need to be close to each other on the network for maximum
application performance.
%
Alternatively, partially or completely replacing the aging and failing
supercomputer with a new system, even if this solution offers less capability,
may be more cost effective in the end.

@@

Supercomputer center policies regarding reliability monitoring, resource
allocation and checkpointing strategies need to be powerful and flexible enough
to facilitate mitigation for such unexpected reliability threats. System
acquisition contracts may need to include performance requirements for degraded
operation, such as a 60\% performance capability if the MTBF drops by a
magnitude.

>>>>

@@

The options of replacing failed components and employing reliability-aware
resource management may not be available or be cost effective for other
supercomputing centers dealing with similar unexpected reliability issues.
%
Replacement components may not be readily available and may have to be
manufactured, which can be impossible for a technology that is no longer supported
by the original manufacturer. Experienced errors or failures may be caused by
software, which can be difficult to fix if the software itself or the deployed
version is no longer supported by the vendor.

There is also a financial aspect, such as a service contract and/or warranty
expiration and the operational budget of a supercomputing center not having
significant funds for replacement/reengineering costs or a manufacturer or
vendor not paying for them.
%
There is additionally the time and financial cost for root-cause analysis and
for developing a hardware and/or software mitigation strategy in the first place.
Fixing a problem requires finding and understanding it, which can be difficult 
in today's complex systems and may require knowledge that only the original 
manufacturer or vendor possesses.

Reliability-aware resource scheduling also has its limitations, as the network
architecture needs to be taken into account. Titan's 3D torus network creates
certainly more challenges for efficient job allocations, than the fat tree
network of Titan's successor, Summit~\cite{olcf:summit}. Nodes associated with
the same job need to be close to each other on the network for maximum
application performance. Also, node outages or less reliable nodes within a
tight group of reliable nodes create resource allocation
holes. Saparating two lower nodes in each cage 0 on Titan would not
create a contiguous partition, yet this group would be the most
reliable in 2016 as was shown in the preceeding section.
%
Alternatively, partially or completely replacing the aging and failing
supercomputer with a new system, even if this solution offers less capability,
may be more cost effective in the end.

@@

Supercomputer center policies regarding reliability monitoring, resource
allocation and checkpointing strategies need to be powerful and flexible enough
to facilitate mitigation for such unexpected reliability threats. System
acquisition contracts may need to include performance requirements for degraded
operation, such as a certain percentage of performance capability if the MTBF
drops by an order of magnitude.
